Spin canonical divisors of log stacky curves

### 2015-06-24 stacky_write_up/spin_cring.tex
major g = 0 revisions, minor typo fixes elsewhere

\begin{cor}\label{cor:genus_0_posiive_delta}
Let $\halfcan$ be a stacky log spin canonical divisor of $(X,\Delta)$ with signature $(g;e_1,\ldots, e_n;\delta)$ such that $2\halfcan=K_\sx+\Delta$ with $\delta=\deg(\Delta)>0$.  Then $\halfcan$ is effective so we can inductively apply lemma \ref{lem:deg1_sat_ind}.
\end{cor}
\begin{proof}
Since $\halfcan$ is a half canonical divisor, 
\[
	2\halfcan=D+\sum \subhalf{e_i}P_i
\] 
where 
\[
	D\sim K_X+\Delta\sim -2\infty+ \Delta
\] is a divisor of $X$ (i.e. with 
possibly negative integer coefficients).  Since $\Delta$ is non-
zero effective, $\deg(D)\ge -1$.   Noting that the coefficient of 
any point $P$ occurring in a stacky divisor of $\sx$ must have 
coefficient lying in $\mathbf{Z}[\frac{1}{e_i}]$ (ranging over $e_i$
s appearing in the characteristic of $(\sx,\Delta)$).

@@

\ssec{Inductive Lemma: Raising Ramification Orders}
\label{ssec:ram_orders}
In this section, we aim to prove the inductive theorem, Theorem
~\ref{thm:ram_order_ind}. Under fairly general
conditions, this theorem will allow us to show that if Theorem ~\ref
{thm:ram_order_ind} holds for a curve with signature
$(0;e_1',\ldots, e_r';0)$ then it also holds for a curve with
signature $(0;e_1,\ldots, e_r;0)$ with $e_i' - e_i \in \{0,2\}$.

\begin{defn}
\label{defn:admissible}
Let $(\sx', 0, L'),(\sx, 0, L)$ be tame, 
separably rooted log stacky spin
curves, both ramified over ramified at $Q_1,\ldots, Q_r$. Suppose
there is a birational map $\sx \rightarrow \sx'$, and let $J \subset
\{1, \ldots, r\}$. Suppose $e_i'+ 2\chi_J(i) = e_i$ where

@@

\begin{thm}
\label{thm:ram_order_ind}
Suppose $(\sx', 0, L')$ \todo{Peter: You should also instantiate $(\sx,0,L)$ as well} is a tame separably rooted log stacky spin
curve and $(\sx', J)$ is admissible with generators $y_{i,e_i'} \in
R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}.
Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}(y_{i,e_i})}{\deg (y_{i,e_i})} \leq \subhalf{e_j'}-
			\frac{1}{\deg(y_{i,e_i})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y,x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z 
			\neq y_{i,e_i},y_{i,e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by
		products of two monomials and that for all $i \in J$, we have $e
		_i > \deg z$ for any generator $z$ of $R'.$ Then any set of
		minimal generators for $I'$ together with any set of relations as in 
		part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

@@

\ssec{Genus Zero Non-effective}
\label{ssec:g_0_non_effective}

\todo{Peter: throughout this section I will provide a few notes of how we can generalize this to arbitrary genus with the appropriate conditions}

In this section, we will prove that if $(\sx , 0, L)$ is a spin curve,
so that $\sx$ has signature $\sigma = (0; e_1, \ldots , e_r; 0)$, then
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1, \ldots,
e_r)$ with relations generated in degree at most $2e$, so long as $\sigma$
does not lie in a finite list of exceptions. This is proven in
Subsection ~\ref{ssec:g_0_noneff_induction}. The proof is a rather
involved induction. In Subsections ~\ref{ssec:g_0_ram_pts},
~\ref{ssec:g_0_ram_orders} \todo{fix} we prove inductive theorems, and in
Subsection ~\ref{ssec:g_0_base} we check the base cases of the
induction. Finally, in Subsection ~\ref{ssec:g_0_exceptional} we
list out the exceptional cases.

@@

\begin{lem}
\label{lem:g_0_admissible_cases}
Let $(\sx', \Delta, \halfcan')$ be a tame log spin curve over a perfect field $k$,
so that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; 0)$.
Then $R' := R(\sx', \Delta, \halfcan')$ is generated as a $k$-algebra by elements
of degree at most $e = \max(5,e_1,\ldots, e_r)$ with relations in degree
at most $2e$ and $(\sx', J)$ is admissible for the following cases:

@@

Using the results of Proposition ~\ref{prop:sat_three_induction}
and Theorem ~\ref{thm:ram_order_ind}, we can inductively raise
the ramification orders and number of ramified points in all of the
cases in Lemma ~\ref{lem:g_0_admissible_cases} to obtain the following
result.

\begin{cor}
\label{cor:g_0_gen_rel_base_cases}
Let $(\sx, \Delta, L)$ be a tame log spin curve over a perfect field $k$, so
that $\sx$ has signature $\sigma = (0; e_1, \ldots, e_r; 0)$. Then the
canonical ring

\[
	R(\sx, 0, L) = \bigoplus_{d = 0}^\infty H^0(\sx, \sco(L)^{\otimes d})
\]

\noindent
is generated as a $k$-algebra by elements of degree at most $e := 
\max(5, e_1, \ldots, e_r)$ with relations in degree at most $2e$, so 
long as $\sigma \notin S$ where

\noindent
is a finite set.
\end{cor}

\begin{proof}
A process of induction raising the ramification orders of
the ramified points of the spin curves corresponding to Lemma
~\ref{lem:g_0_admissible_cases} cases (a)-(c) via Theorem
~\ref{thm:ram_order_ind} will demonstrate the desired generator and
relation degree results for curves with three ramified points.
Similarly, this induction process on Lemma
~\ref{lem:g_0_admissible_cases} case (d) will give the desired
result for curves with four ramified points excluding signature
$(0; 3, 3, 3, 3; 0)$. A process of induction increasing the
number of ramified points via Proposition
~\ref{prop:sat_three_induction} on Lemma ~\ref{lem:g_0_admissible_cases}
case (e) combined with the induction process on ramification orders
will yield the result for curves with at least five ramified points.
\todo{Detailed description of how the inductive lemmata yield e and 2e}
\end{proof}

\ssec{Exceptional Cases}
\label{ssec:g_0_exceptional}
Here we present the explicit generators and relations of the remaining
cases given by signatures in the finite set

@@

>>>>

\begin{cor}\label{cor:genus_0_positive_delta}
Let $\halfcan$ be a stacky log spin canonical divisor of $(X,\Delta)$ with signature $(g;e_1,\ldots, e_n;\delta)$ such that $2\halfcan=K_\sx+\Delta$ with $\delta=\deg(\Delta)>0$.  Then $\halfcan$ is effective so we can inductively apply lemma \ref{lem:deg1_sat_ind}.
\end{cor}
\begin{proof}
Since $\halfcan$ is a half canonical divisor, 
\[
	2\halfcan=D+\sum \subhalf{e_i}P_i
\] 
where 
\[
	D\sim K_X+\Delta\sim -2\infty+ \Delta
\] is a divisor of $X$ (i.e. with 
possibly negative integer coefficients).  Since $\Delta$ is non-
zero effective, $\deg(D)\ge -1$.   Noting that the coefficient of 
any point $P$ occurring in a stacky divisor of $\sx$ must have 
coefficient lying in $\mathbf{Z}[\frac{1}{e_i}]$ (ranging over $e_i$
s appearing in the characteristic of $(\sx,\Delta)$).

@@

\ssec{Inductive Lemma: Raising Ramification Orders}
\label{ssec:ram_orders}
In this section, we aim to prove the inductive theorem, Theorem
~\ref{thm:ram_order_ind}. Under fairly general
conditions, this theorem will allow us to show that if Theorem ~\ref
{thm:ram_order_ind} holds for a curve with signature
$(0;e_1',\ldots, e_r';0)$ then it also holds for a curve with
signature $(0;e_1, \ldots, e_r; 0)$ with $e_i' - e_i \in \{0,2\}$.

\begin{defn}
\label{defn:admissible}
Let $(\sx', 0, L'),(\sx, 0, L)$ be tame, 
separably rooted log stacky spin
curves, both ramified over ramified at $Q_1,\ldots, Q_r$. Suppose
there is a birational map $\sx \rightarrow \sx'$, and let $J \subset
\{1, \ldots, r\}$. Suppose $e_i'+ 2 \chi_J (i) = e_i$ where

@@

\begin{thm}
\label{thm:ram_order_ind}
Suppose $(\sx', 0, L')$ \todo{Peter: You should also instantiate $(\sx, 0, L)$ as well} is a tame separably rooted log stacky spin
curve and $(\sx', J)$ is admissible with generators $y_{i,e_i'} \in
R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}.
Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in
		H^0(\sx, e_i(K_\sx))$ so that
		\begin{align*}
			-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
		\end{align*}
		and
		\begin{align*}
			\frac{-\ord_{Q_j}(y_{i,e_i})}{\deg (y_{i,e_i})} \leq \subhalf{e_j'}-
			\frac{1}{\deg(y_{i,e_i})e_j'}
		\end{align*}
		for all $j \in J$ with $j \neq i$.
	\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b$ with $a \geq 0,
		b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally
		generate $R$ over $R'$.
	\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with
		any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with
		block order. Let $R = k[y,x]/I.$ Then,
		\begin{align*}
			in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z 
			\neq y_{i,e_i},y_{i,e_i'} \rangle
		\end{align*}
		as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by
		products of two monomials and that for all $i \in J$, we have $e
		_i > \deg z$ for any generator $z$ of $R'.$ Then any set of
		minimal generators for $I'$ together with any set of relations as in 
		part (c) minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}

@@

\ssec{Genus Zero Non-effective}
\label{ssec:g_0_non_effective}
In this section, we will prove that if $(\sx , 0, L)$ is a spin curve,
so that $\sx$ has signature $\sigma = (0; e_1, \ldots , e_r; 0)$, then
$R(\sx , 0, L)$ is generated in degree at most $e := \max(5, e_1, \ldots,
e_r)$ with relations generated in degree at most $2e$, so long as $\sigma$
does not lie in a finite list of exceptions. This is proven in
Subsection ~\ref{ssec:g_0_noneff_induction}. The proof is a rather
involved induction. In Subsections ~\ref{ssec:g_0_ram_pts},
~\ref{ssec:g_0_ram_orders} \todo{fix} we prove inductive theorems, and in
Subsection ~\ref{ssec:g_0_base} we check the base cases of the
induction. Finally, in Subsection ~\ref{ssec:g_0_exceptional} we
list out the exceptional cases.

@@

\begin{lem}
\label{lem:g_0_admissible_cases}
Let $(\sx', \Delta, \halfcan')$ be a tame log spin curve over a
perfect field $k$, so that $\sx$ has signature $\sigma = (0; e_1,
\ldots, e_r; 0)$. Then $R' := R(\sx', \Delta, \halfcan')$ has a
presentation

is
generated as a $k$-algebra by elements of degree at most $e = \max(5
,e_1,\ldots, e_r)$ with relations in degree at most $2e$ and $(\sx
', J)$ is admissible for the following cases:

@@

Through the application of two processes of induction, by adding 
points via Proposition ~\ref{prop:sat_three_induction} and by raising
ramification orders via Theorem ~\ref{thm:ram_order_ind},
we obtain Theorem ~\ref{thm:g_0_generators_relations} for any tame
log spin curve $(\sx, \Delta, L)$ over a perfect field $k$ with
signature $\sigma$ covered by this induction process.
In particular, this induction process covers all signatures $\sigma
= (0; e_1, \ldots, e_r; 0)$ such that either $r \geq 5$ or
for some listed base case $(\sigma', J)$ with $\sigma' = (0; e_1',
\ldots, e_r'; 0)$, $e_i \geq e_i'$ for all $i$.
Before we prove the main theorem, we describe the cases not be
covered by induction.

\ssec{Exceptional Cases}
\label{ssec:g_0_exceptional}
First, we describe the cases that would not be covered by induction.
Here we present the explicit generators and relations of the remaining
cases given by signatures in the finite set

@@

\begin{proof}

%By Lemma ~\ref{lem:g_0_admissible_cases}, we know that for
%any tame log spin curves $(\sx', \Delta, \halfcan')$ over a perfect
%field $k$ with signature $\sigma'$ and set $J$ such that $(\sigma',
%J)$ is one of the listed base cases, $(\sigma', J)$ is admissible.
%Furthermore, $R' := R(\sx', \Delta, \halfcan')$ is generated as a $k
%$- algebra by elements of degree at most $e = \max(5, e_1, \ldots, e
%_r)$ with relations in degree at most $2e$. It also admits a
%presentation
%
%\begin{align*}
	%R' \cong \left( k[x_1, \ldots, x_m] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
%\end{align*}
%
%\noindent
%given by admissibility satisfying the statements that for all $i \in 
%J$, $e_i > \deg z$ for any generator $z$ of $R'$ and that $in_\prec
%(I')$ is minimally generated by products of two monomials under
%some ordering.
%
%Then we can apply two processes of induction, by adding points via
%Proposition ~\ref{prop:sat_three_induction} and by raising
%ramification orders via Theorem ~\ref{thm:ram_order_ind},
%to obtain the desired result for any tame log spin curve $(\sx,
%\Delta, L)$ over a perfect field $k$ with signature $\sigma$
%covered by this induction process.
%
%
%
%A process of induction raising the ramification orders of
%the ramified points of the spin curves corresponding to Lemma
%~\ref{lem:g_0_admissible_cases} cases (a)-(c) via Theorem
%~\ref{thm:ram_order_ind} will demonstrate the desired generator and
%relation degree results for curves with three ramified points.
%Similarly, this induction process on Lemma
%~\ref{lem:g_0_admissible_cases} case (d) will give the desired
%result for curves with four ramified points excluding signature
%$(0; 3, 3, 3, 3; 0)$. A process of induction increasing the
%number of ramified points via Proposition
%~\ref{prop:sat_three_induction} on Lemma ~\ref{lem:g_0_admissible_cases}
%case (e) combined with the induction process on ramification orders
%will yield the result for curves with at least five ramified points.
\end{proof}
